Video Games Sales Analysis by Bayan AlArifi
========================================================

> This report explores a dataset generated by a scrape of vgchartz.com. The scraping script is available at https://github.com/GregorUT/vgchartzScrape. The dataset contains video games sales and attributes for approximately 11493 video games. The dataset contains 16,598 observations and 11 variables.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(dplyr) # a package for dataframe manipulation
library(ggplot2) # a package for creating plots
library(corrplot) # a package for graphical display of a correlation matrix, confidence
library(tidyr)  # a package for converting data to a long format
library(PerformanceAnalytics) # a package used to produce a correlation matrix
library(gridExtra) # a package for displaying plots in a grid side-by-side
library(reshape2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
#getwd() # get current directory
dataset <- read.csv('vgsales.csv')
```

These variables are:
```{r echo=FALSE, message=FALSE, warning=FALSE, Data_Columns}
# List the columns/variables in the dataset
names(dataset)
```

A summary of various model fitting functions' results for each of the variables/features:
```{r echo=FALSE, message=FALSE, warning=FALSE, Data_Functions}
# Get a summary of various model fitting functions' results
summary(dataset)
```

Checking the first couple of rows..
```{r echo=FALSE, message=FALSE, warning=FALSE, Data_Content}
head(dataset)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# define theme variables to be used in the plots
text <- element_text(size=7)
key <- unit(1,"line")
title <- element_text(size=9)
axis <- element_text(angle = 90, hjust = 1)

theme <-  theme(legend.text=text, legend.key.size = key, legend.title=title, legend.position = "top", axis.text.x = axis)
theme1 <- theme(legend.position = "none")
theme2 <- theme(legend.position="top", axis.text.x = axis)
theme3 <- theme(axis.text.x = axis)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# define plot specification variables to be used in the plots
plot_specs <- geom_bar(stat = "identity")
plot_specs2 <- geom_bar(stat = "identity", position = "dodge")
spec <- ylab("Global Sales in Millions") +
        theme +
        coord_flip()
```

# Univariate Plots Section

The feature/variable Rank is the rank of the 16327 video games in the dataset considering their global sales, where rank #1 is the top selling. Thus, there was no need to plot it here.

**Global Sales**

Since 1980, video games global sales reached a total of 8920.44 million.
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(dataset$Global_Sales)
sum(dataset$Global_Sales)
```

**Yearly Releases**
```{r echo=FALSE, message=FALSE, warning=FALSE}
dataset <- dataset[dataset$Year!='N/A', ] # remove N/A rows
summary(dataset$Year)
```

Most video games were released in 2008 (1428 games) and 2009 (1431 games).

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot video game releases per year
ggplot(dataset, aes(Year)) +
  theme3 +
  labs(title='Video Games Releases per Year') +
  geom_bar(fill = '#F57670')
```

**Video Games Releases**
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(dataset$Name)
```

The top released video game is 'Need for Speed: Most Wanted'.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Find the top 5 released video games
top <- table(dataset$Name)
top <- sort(top, decreasing=TRUE)[1:5]
top <- as.data.frame(top)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot top 5 released video games
ggplot(top, aes(x=Var1, y=Freq)) +
  xlab('Top 5 Video Games') +
  ylab('Count') +
  labs(title='Top 5 Released Video Games') +
  geom_bar(fill = '#6bbd80', stat = "identity")
```

**Platforms**
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(dataset$Platform)
```

The top releasing video game platforms are PS2 (2127) and DS (2133), out of 31 total platforms.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Find the top 5 released video game platforms
top <- table(dataset$Platform)
top <- sort(top, decreasing=TRUE)[1:5]
top <- as.data.frame(top)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot top 5 released video game platforms
ggplot(top, aes(x=Var1, y=Freq)) +
  xlab('Top 5 Platforms') +
  ylab('Count') +
  labs(title='Top 5 Releasing Platforms') +
  geom_bar(fill = '#e570f1', stat = "identity")
```

**Publishers**
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(dataset$Publisher)
```

The top video games publisher is Electronic Arts (1339), out of a total of 579 publishers.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Find the top 5 releasing video game publishers
top <- table(dataset$Publisher)
top <- sort(top, decreasing=TRUE)[1:5]
top <- as.data.frame(top)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot top 5 releasing video game publishers
ggplot(top, aes(x=Var1, y=Freq)) +
  xlab('Top 5 Publishers') +
  ylab('Count') +
  labs(title='Top 5 Releasing Publishers') +
  geom_bar(fill = '#63aef3', stat = "identity")
```

**Genres**
```{r echo=FALSE, message=FALSE, warning=FALSE}
summary(dataset$Genre)
```

The top selling video games genre is Action with a total of 3253 video game releases, out of a total of 12 genres. Action Games contribute to almost 20% of the games released. Sports games coming in second with 14% (2304 releases). Overall all top 5 genres contribute to 61% of the games released.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# find the top releasing genres
top <- table(dataset$Genre)
top <- sort(top, decreasing=TRUE)
top <- as.data.frame(top)
total <- 16598
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot video game genres
ggplot(top, aes(x="", y=Freq, fill=Var1)) + geom_bar(stat="identity", width=1) +
  coord_polar("y", start=0) + geom_text(aes(label = paste0(round((top$Freq*100)/total), "%")), position = position_stack(vjust = 0.5)) +
  labs(x = NULL, y = NULL, fill = NULL, title = "Video Games Genres") +
  theme_classic() +
  theme(axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), plot.title = element_text(hjust = 0.5, color = "#666666"))
```

# Univariate Analysis

### What is the structure of your dataset?
There are 16,598 observations about 11493 video games, with 11 features, as descrived below.

**Variables Dictionary:**

* Rank - Ranking of overall sales

* Name - The games name

* Platform - Platform of the games release (i.e. PC,PS4, etc.)

* Year - Year of the game's release

* Genre - Genre of the game

* Publisher - Publisher of the game

* NA_Sales - Sales in North America (in millions)

* EU_Sales - Sales in Europe (in millions)

* JP_Sales - Sales in Japan (in millions)

* Other_Sales - Sales in the rest of the world (in millions)

* Global_Sales - Total worldwide sales.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Checking the data structure, variables, and types
str(dataset)
```

### What is/are the main feature(s) of interest in your dataset?
The main features in the data set are Global_Sales and Year, and discovering the trend around them, i.e. what are the top globally selling value of the other features, and top other features values over time.

### What other features in the dataset do you think will help support your \ investigation into your feature(s) of interest?
The month of the year could likely contribute to the rise of sales. For example, certain video game genres could be always popular in a certain month of the year. It would be interesting to see how the other regions around the world's sales do, rather than just having the variable Other_Sales. Lastly, demographic data on the buyers of these games could tell us a lot on why certain feature values happen to be more popular than others. In fact, such kind of data can also help us predict the future sales considering the buyers demographic and behaviour.

### Did you create any new variables from existing variables in the dataset?
In order to create the regional sales comparison plots in the **Multivariate** section, I needed to combine the regional sales columns (NA_Sales, EU_Sales, JP_Sales, Other_Sales, Global_Sales) by using the gather function from the tidyr package.

### Of the features you investigated, were there any unusual distributions? \ Did you perform any operations on the data to tidy, adjust, or change the form \ of the data? If so, why did you do this?
I had to drop observations where the Year had an N/A value.

# Bivariate Plots Section

```{r echo=FALSE}
# # plot global sales per year
# ggplot(dataset, aes(Year, Genre, fill = Genre)) +
#   geom_bar(stat = "identity") +
#   ggtitle("Genre Distribution Over the Years") +
#   xlab("Year") +
#   ylab("Genre") +
#   theme3
```

```{r echo=FALSE}
# # plot global sales per year
# ggplot(dataset, aes(Year, Platform, fill = Platform)) +
#   geom_bar(stat = "identity") +
#   ggtitle("Platform Distribution Over the Years") +
#   xlab("Year") +
#   ylab("Platform") +
#   theme3
```

**Global Sales per Year**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Find the global sales per year
top_year <- dataset %>%
             group_by(Year) %>%
             summarize(Sales = sum(Global_Sales))

summary(top_year$Sales)
```

The highest global sales were in 2008, with a total of 678.90 million. 2009 comes in second with slightly less sales. On average, the yearly global sales reach 226.16 million as shown in the purple dashed line.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# plot global sales per year
ggplot(top_year, aes(Year, Sales, fill = Sales)) + 
  geom_bar(stat = "identity") +
  ggtitle("Global Sales per Year") +
  ylab("Global Sales in Millions") +
  xlab("Year") +
  geom_hline(yintercept = mean(top_year$Sales), linetype="dashed", color = "Purple") +
  theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1))
```

**Top Globally Selling Genres**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# find the top 5 globally selling video games
top_genres <- dataset %>%
             group_by(Genre) %>%
             summarize(Sales = sum(NA_Sales), Percentage = Sales/sum(dataset$Global_Sales) * 100) %>%
             arrange(desc(Sales)) %>%
             head(5)

summary(top_genres$Sales)
```

The top globally selling video game genre is Action with 1722.9 million. On average, the global sales reach 591.1 million as shown in the black dashed line.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# plot the top 5 globally selling genres
ggplot(top_genres, aes(reorder(Genre, Sales), Sales, fill = Genre)) + 
  geom_bar(stat = "identity", fill = '#F57670') +
  ggtitle("Top 5 Globally Selling Genres") +
  xlab("Genre") +
  geom_hline(yintercept = mean(top_genres$Sales), linetype="dashed", color = "Black") +
  spec
```

**Top Globally Selling Publishers**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# find the top 5 globally selling publishers
top_publishers <- dataset %>%
             group_by(Publisher) %>%
             summarize(Sales = sum(Global_Sales), Percentage = Sales/sum(dataset$Global_Sales) * 100) %>%
             arrange(desc(Sales)) %>%
             head(5)

summary(top_publishers$Sales)
```

The top globally selling video game publisher is Nintendo with 1784.4 million. On average, the global sales reach 936.0 million as shown in the black dashed line.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# plot the top 5 globally selling genres
ggplot(top_publishers, aes(reorder(Publisher, Sales), Sales, fill = Publisher)) + 
  geom_bar(stat = "identity", fill = '#6bbd80') +
  ggtitle("Top 5 Globally Selling Publishers") +
  xlab("Publisher") +
  geom_hline(yintercept = mean(top_publishers$Sales), linetype="dashed", color = "Black") +
  spec
```

**Top Globally Selling Platforms**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# find the top 5 globally selling platforms
top_platforms <- dataset %>%
             group_by(Platform) %>%
             summarize(Sales = sum(Global_Sales), Percentage = Sales/sum(dataset$Global_Sales) * 100) %>%
             arrange(desc(Sales)) %>%
             head(5)

summary(top_platforms$Sales)
```

The top globally selling video game platforms is PS2 with 988.5 million. On average, the global sales reach 936.0 million as shown in the black dashed line.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# plot the top 5 globally selling platforms
ggplot(top_platforms, aes(reorder(Platform, Sales), Sales, fill = Platform)) + 
  geom_bar(stat = "identity", fill = '#e570f1') +
  ggtitle("Top 5 Globally Selling Platforms") +
  xlab("Platform") +
  geom_hline(yintercept = mean(top_publishers$Sales), linetype="dashed", color = "Black") +
  spec
```

**Top Globally Selling Video Games**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# find the top 5 globally selling video games
top_games <- dataset %>%
             group_by(Name) %>%
             summarize(Sales = sum(Global_Sales), Percentage = Sales/sum(dataset$Global_Sales) * 100) %>%
             arrange(desc(Sales)) %>%
             head(5)

summary(top_games$Sales)
```

The top globally selling video game is Wii Sport with 82.74 million. On average, the global sales reach 51.13 million as shown in the black dashed line.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# plot the top 5 globally selling video games
ggplot(top_games, aes(reorder(Name, Sales), Sales, fill = Name)) + 
  geom_bar(stat = "identity", fill = '#63aef3') +
  ggtitle("Top 5 Globally Selling Video Games") +
  xlab("Video Game") +
  geom_hline(yintercept = mean(top_games$Sales), linetype="dashed", color = "Black") +
  spec
```

**Regional Sales**

The distribution of each regional sales variable is shown on the diagonal of the below plot. On the bottom of the diagonal, the bivariate scatter plots with a fitted line are displayed. On the top of the diagonal, the value of the correlation plus the significance level as stars. Each significance level is associated with a symbol:
p-values(0, 0.001, 0.01, 0.05, 0.1, 1) <=> symbols(“***”, “**”, “*”, “.”, " “).
As we can see in the matrix, NA_Sales (0.94) and EU_Sales (0.90) indicate an almost perfect positive correlation with Global_Sales. Moreover, if we look into the behavious of the NA_Salesand EU_Sales, they seem to move in the same direction. In general, all of the correlation coefficient values are greater than zero. Thus, there is a positive relationship between these variables.

```{r echo=FALSE, message=FALSE, warning=FALSE}
sales <- select(dataset, NA_Sales, EU_Sales, JP_Sales, Other_Sales, Global_Sales)
chart.Correlation(sales, histogram=FALSE, pch=19, title='Regional Sales Correlation Matrix')
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \ investigation. How did the feature(s) of interest vary with other features in \ the dataset?
2008 and 2009 were the most successful years for the game community. Action games are more famous nowadays. Nintendo and PS2 are behind the highest global sales. The top globally selling video game since 1980 is Wii Sports.

### Did you observe any interesting relationships between the other features \ (not the main feature(s) of interest)?
No, not in my bivariate analysis.

### What was the strongest relationship you found?
The one explained in the correlation matrix in the regional sales section.

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots}
# a function for creating the top dataset in this section
findTop <- function(feature){
  # per year
  if (feature == 'Genre1') {
    top <- dataset %>%
           group_by(Year, Genre)
  } else if (feature == 'Platform1') {
    top <- dataset %>%
           group_by(Year, Platform)
  } else if (feature == 'Publisher1') {
    top <- dataset %>%
           group_by(Year, Publisher)
  } else if (feature == 'Name1') {
    top <- dataset %>% 
           group_by(Year, Name)
  }
  # per platform
  else if (feature == 'Genre2') {
    top <- dataset %>%
           group_by(Genre, Name)
  } else {
    top <- dataset %>%
           group_by(Platform, Name)
  }

  top <- top %>%
         summarize(Sales = sum(Global_Sales)) %>%
         top_n(1)
  return(top)
}
```

**Globally Selling Genres per Year**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# find the top genres in each year
top_genre <- findTop('Genre1')
summary(top_genre$Sales)
```

Action is the top globally selling genre since 1980 with 139.36 million in 2009. Sports comes second with slightly less sales in 2006. Action also happens to be the top globally selling genre in 17 years (1981, 1986, 2001-2005, 2007-2016). On a yearly average, top video game publisher sales reach 53.43 million.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# plot top globally selling genres per year
ggplot(top_genre, aes(Year, Sales, fill = Genre)) + 
  plot_specs +
  ggtitle("Top Globally Selling Genres per Year") +
  ylab("Global Sales in Millions") +
  theme
```

**Globally Selling Publishers per Year**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# find the top publishers in each year
top_publisher <- findTop('Publisher1')
summary(top_publisher$Sales)
```

Nintendo is the top globally selling publishers since 1980 with 205.61 million in 2006. It also happens to be the top globally selling publisher in 24 years (1983-1994, 1996, 1998-2001, 2005-2009, 2012, 2014). Electronic Arts comes second. On a yearly average, top video game publisher sales reach 50.57 million.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# plot top globally selling publishers per year
ggplot(top_publisher, aes(Year, Sales, fill = Publisher)) + 
  plot_specs +
  ggtitle("Top Globally Selling Publishers per Year") +
  ylab("Global Sales in Millions") +
  theme
```

**Globally Selling Platfroms per Year**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# find the top platforms in each year
top_platform <- findTop('Platform1')
summary(top_platform$Sales)
``` 

PS2 is the top globally selling platform since 1980 with 211.78 million in 2004. Wii comes second with slightly less sales in 2009. However, NES happens to be the top globally selling platforms in 6 years (1995-2000), and PS2 the top in 5 years (2001-2005). On a yearly average, top video game publisher sales reach 92.34 million.

```{r echo=FALSE, message=FALSE, warning=FALSE}           
# plot top globally selling platforms per year
ggplot(top_platform, aes(Year, Sales, fill = Platform)) + 
  plot_specs +
  ggtitle("Top Globally Selling Platform per Year") +
  ylab("Global Sales in Millions") +
  theme
```

**Globally Selling Games per Year**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# find the top platforms in each year
top_games <- findTop('Name1')
summary(top_games$Sales)
``` 

Wii Sports is the top globally selling game since 1980 with 82.74 million in 2006. Super Mario Bros. comes second in 1985. On a yearly average, top video games sales reach 18.93 million.

```{r echo=FALSE, message=FALSE, warning=FALSE}           
# plot top globally selling games per year
ggplot(top_games, aes(Year, Sales, fill = Name)) + 
  plot_specs +
  ggtitle("Top Globally Selling Games per Year") +
  ylab("Global Sales in Millions") +
  theme
```

**Globally Selling Video Games per Platform**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# find the top games in each platform
top_platform_game <- findTop('Platform2')
summary(top_platform_game$Sales)
``` 

```{r echo=FALSE, message=FALSE, warning=FALSE}           
# plot top globally selling games per genre
ggplot(top_platform_game, aes(Platform, Sales, fill = Name)) + 
  plot_specs +
  ggtitle("Top Globally Selling Video Games per Platform") +
  ylab("Global Sales in Millions") +
  theme
```

**Globally Selling Video Games per Genre**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# find the top games in each genre
top_genre_game <- findTop('Genre2')
summary(top_genre_game$Sales)
``` 

```{r echo=FALSE, message=FALSE, warning=FALSE}           
# plot top globally selling games per genre
ggplot(top_genre_game, aes(Genre, Sales, fill = Name)) + 
  plot_specs +
  ggtitle("Top Globally Selling Video Games per Genre") +
  ylab("Global Sales in Millions") +
  theme
```

**Global and Regional Sales**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Combine regional sales columns
regional_sales <- select(dataset, Year, NA_Sales, EU_Sales, JP_Sales, Other_Sales, Global_Sales)
regional_sales <- gather(regional_sales, key=Region, value=Sales, NA_Sales, EU_Sales, JP_Sales, Other_Sales, Global_Sales)

summary(dataset$Global_Sales)
```

The below two plots illustrate a comparision of the regional and global sales. Follows that plots that show each region's sales compared to the global ones over time.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# plot the yearly regional sales comparision
ggplot(dataset, aes(Year)) +
  theme2 +
  labs(title='Video Games Yearly Regional Sales', colour = "Region") +
  ylab("Sales in Millions") +
  geom_line(aes(y = Global_Sales, colour = "Globe")) + 
  geom_line(aes(y = NA_Sales, colour = "North America")) +
  geom_line(aes(y = EU_Sales, colour = "Europe")) +
  geom_line(aes(y = JP_Sales, colour = "Japan")) +
  geom_line(aes(y = Other_Sales, colour = "Other"))
```

There is a huge fall in sales in the last 5-6 years, escpecially in the North American region.
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Line plot the yearly regional sales comparision
ggplot(regional_sales, aes(Year, Sales, group=Region)) +
  geom_line(aes(color=Region)) +
  labs(title='Video Games Yearly Regional Sales', colour = "Region") +
  ylab("Sales in Millions") +
  #geom_point(aes(color=Region))+
  theme2
```

**Global and North America Yearly Sales**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Combine regional sales columns
regional_sales <- select(dataset, Year, NA_Sales, Global_Sales)
summary(regional_sales$NA_Sales)
regional_sales <- gather(regional_sales, Region, Sales, NA_Sales:Global_Sales)
```

The NA sales reached a peak of 41.4900 million in 2006. On average, their sales are 0.2654 million.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot NA and Global yearly sales
ggplot(regional_sales, aes(Year, Sales, fill=Region)) +
  theme2 +
  labs(title='North American and Global Yearly Video Games Sales', colour = "Region") +
  ylab("Sales in Millions") +
  plot_specs2
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
# # Line plot NA and Global yearly sales
# ggplot(regional_sales, aes(Year, Sales, group=Region)) +
#   geom_line(aes(color=Region)) +
#   #geom_point(aes(color=Region))+
#   theme2
```

**Global and European Yearly Sales**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Combine regional sales columns
regional_sales <- select(dataset, Year, EU_Sales, Global_Sales)
summary(regional_sales$EU_Sales)
regional_sales <- gather(regional_sales, Region, Sales, EU_Sales:Global_Sales)
```

The EU sales reached a peak of 29.0200 million in 2006. On average, their sales are 0.1476 million.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot European and Global yearly sales
ggplot(regional_sales, aes(Year, Sales, fill=Region)) +
  theme2 +
  labs(title='European and Global Yearly Video Games Sales', colour = "Region") +
  ylab("Sales in Millions") +
  plot_specs2
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
# # Line plot European and Global yearly sales
# ggplot(regional_sales, aes(Year, Sales, group=Region)) +
#   geom_line(aes(color=Region)) +
#   #geom_point(aes(color=Region))+
#   theme2
```

**Global and Japanese YearlySales**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Combine regional sales columns
regional_sales <- select(dataset, Year, JP_Sales, Global_Sales)
summary(regional_sales$JP_Sales)
regional_sales <- gather(regional_sales, Region, Sales, JP_Sales:Global_Sales)
```

The JP sales reached a peak of 10.22000 million in 1996. On average, their sales are 0.07866 million.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot Japanese and Global yearly sales
ggplot(regional_sales, aes(Year, Sales, fill=Region)) +
  theme2 +
  labs(title='Japanese and Global Yearly Video Games Sales', colour = "Region") +
  ylab("Sales in Millions") +
  plot_specs2
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# # Line plot Japanese and Global yearly sales
# ggplot(regional_sales, aes(Year, Sales, group=Region)) +
#   geom_line(aes(color=Region)) +
#   #geom_point(aes(color=Region))+
#   theme2
```

**Global and Other Yearly Sales**
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Combine regional sales columns
regional_sales <- select(dataset, Year, Other_Sales, Global_Sales)
summary(regional_sales$Other_Sales)
regional_sales <- gather(regional_sales, Region, Sales, Other_Sales:Global_Sales)
```

The other regional sales reached a peak of 10.57000 million in 2004. On average, their sales are 0.04832 million.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Bar plot Other and Global yearly sales
ggplot(regional_sales, aes(Year, Sales, fill=Region)) +
  theme2 +
  labs(title='Other and Global Yearly Video Games Sales', colour = "Region") +
  ylab("Sales in Millions") +
  plot_specs2
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
# # Line plot Other and Global yearly sales
# ggplot(regional_sales, aes(Year, Sales, group=Region)) +
#   geom_line(aes(color=Region)) +
#   #geom_point(aes(color=Region))+
#   theme2
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \ investigation. Were there features that strengthened each other in terms of \ looking at your feature(s) of interest?
Although the action genre was the popular famous since 2000, the second popular genre Sports (as we saw in the bivariat analysis) peaked in 2006. This was further confirmed after plotting A. the globally selling games per year and finding out it was Wii Sports from the Sports genre in 2006, and B. the globally selling games per genre, and C. the globally selling games platform which of course is Wii.

### Were there any interesting or surprising interactions between features?
The global and regional yearly sales for North America and Europe further confirm the results found in the correlation matrix, where the plots show that a rise in each of these regions whenever the global sales rises. On another note, the globally selling per year plots for publishers and platforms show an interesting pattern, where for example between 2006 and 2009 the Nintendo publisher was dominant, and the Wii platform was dominant. This makes sense since Nintendo owns Wii, and was released for the first time in November 19, 2006. Video game publishers that also happen to own platforms/consoles publish a couple of games that work with their new product to help market it. Other publishers follow to ensure a competitor's advantage.

------

# Final Plots and Summary

### Plot One: **Globally Selling Platfroms per Year**
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
# Find the top platforms in each year
top_platform <- dataset %>%
             group_by(Year, Platform) %>%
             summarize(Sales = sum(Global_Sales)) %>%
             arrange(desc(Sales)) %>%
             top_n(1)
``` 

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot top globally selling platforms per year
ggplot(top_platform, aes(Year, Sales, fill = Platform)) + 
  geom_bar(stat = "identity") +
  ggtitle("Top Globally Selling Platform per Year") +
  ylab("Global Sales in Millions") +
  theme
``` 

### Description One
PS2 is the top globally selling platform since 1980 with 211.78 million in 2004. Wii comes second with slightly less sales in 2009. However, NES happens to be the top globally selling platforms in 6 years (1995-2000), and PS2 the top in 5 years (2001-2005). On a yearly average, top video game publisher sales reach 92.34 million. What is interesting about this plot is that it shows a pattern where a top selling publisher continues to hold that rank for a couple of years (3-6 years), with only one exception in the X360 case in 2010. Possible reasons for this could be that Wii did not have many popular games at the time.

### Plot Two: **Top Globally Selling and Releasing Publishers**
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
# Find the top 5 releasing video game publishers
top <- table(dataset$Publisher)
top <- sort(top, decreasing=TRUE)[1:5]
top <- as.data.frame(top)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot top 5 releasing video game publishers
plot1 = ggplot(top, aes(x=Var1, y=Freq)) +
        xlab('Top 5 Publishers') +
        ylab('Releases') +
        labs(title='Top 5 Releasing Publishers') +
        geom_bar(fill = '#63aef3', stat = "identity")
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Find the top 5 globally selling genres
top_publishers <- dataset %>%
             group_by(Publisher) %>%
             summarize(Sales = sum(Global_Sales), Percentage = Sales/sum(dataset$Global_Sales) * 100) %>%
             arrange(desc(Sales)) %>%
             head(5)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot the top 5 globally selling genres
plot2 = ggplot(top_publishers, aes(reorder(Publisher, Sales), Sales, fill = Publisher)) + 
        geom_bar(stat = "identity", fill = '#e570f1') +
        ggtitle("Top 5 Globally Selling Publishers") +
        ylab("Global Sales in Millions") +
        xlab("Publisher") +
        theme1

grid.arrange(plot1, plot2, nrow=2)
```

### Description Two
The top releasing publisher is Electronic Arts with 1339 releases. Also in the top 5 are Activision, Namco Bandai Games, Ubisoft, and Konami digital Entertainment. However, the top globally selling publisher is Nintendo with 1784.4 million, whcih doesn't happen to be in any of the top 5 releasing publishers. Same thing goes for Sony Computer Entertainment and Ubisoft. This goes to show that releasing to many video games does not garantee more sales.

### Plot Three: **Yearly Global Sales and Releases**
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
# Plot video game releases per year
plot1 = ggplot(dataset, aes(Year)) +
        theme3 +
        ylab("Releases") +
        labs(title='Video Games Releases per Year') +
        geom_bar(fill = '#F57670')
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Find the top global sales per year
top_year <- dataset %>%
             group_by(Year) %>%
             summarize(Sales = sum(Global_Sales))
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot global sales per year
plot2 = ggplot(top_year, aes(Year, Sales, fill = Sales)) + 
        geom_bar(stat = "identity") +
        ggtitle("Global Sales per Year") +
        ylab("Sales in Millions") +
        xlab("Year") +
        theme(legend.position = "none", axis.text.x = element_text(angle = 90, hjust = 1))

grid.arrange(plot1, plot2, nrow=2)
```
### Description Three
Most video games were released in 2008 (1428 games) and 2009 (1431 games). The highest global sales were also in 2008, with a total of 678.90 million. 2009 comes in second with slightly less sales. Since 1996, the video game industry flourished. It continued to grow in the number of game releases, attracting more global sales, peaking in 2008-2009 as we can see in both plots of global sales/releases per year. We can also notice from the two plots that after the peak it started to gradually decrease.

------

# Reflection

* Since 1980, video games global sales reached a total of 8920.44 million.

* Action games have been the most popular for over a decade. They contribute to almost 20% of the games released.

* Play Station is the most popular platform.

* Wii Sports is the top video game by revenue.

* Namco Bandai Company is producing maximum number of successful games.

* There is huge spike in the number of releases after 2000 and it peaked during 2008 and 2009.

* 2009 was the most successful year for the video games community.

* The last 5-6 years has seen a decrease in the revenue.

* EA is the top most publisher by number of releases.

* DS is the top most platform by number of releases.

* Wii Sports in 2006 generated huge revenue compared to all other games along the years.

* Nintendo has been dominating the market as the top publisher by revenue in 24 years since 1980.

* Releasing too many video games does not garantee more sales as we have seen in the case of Nintendo, Sony Computer Entertainment and Ubisoft, where they were in the top selling publishers list but not in the top releasing publishers list.

* Once a platform clicks in the market i.e. top selling, it goes on to rule for few years (3-6 years).

* North America and Europe contribute 75% of the market share by revenue, and both had their sales peak in 2006, although the global sales peak was in 2008.

One aspect that limited this analysis is the absence of the sales of other parts of the world, rather than just having the variable Other_Sales. Also, the month of the year could likely contribute to the rise of sales. For example, certain video game genres could be always popular in a certain month of the year. For future work, I would like to find data on the buyers demographic which could explain a lot of the resons behind the trends found, and improve the analysis. It will also help draw recommendations to the decision makers in this industry on where to focus their budget; which genre, which region, which platform, etc.